Cancer Applications of Choropleth Maps, and the Potential of Cartograms and Alternative Map Displays

### 2019-12-18 review_article.Rmd

---
title: Cancer Applications of Choropleth Maps, and the Potential of Cartograms and
  Alternative Map Displays
author: Stephanie Kobakian, Jessie Roberts and Dianne Cook
csl: springer-vancouver.csl
output:
  word_document:
    toc: no
    toc_depth: '3'
  pdf_document:
    dev: jpeg
    fig_caption: yes
    number_sections: yes
    toc: no
    toc_depth: 3
geometry: margin = 1.2in
header-includes:
- \usepackage{setspace}\doublespacing
- \usepackage{fancyhdr}\pagestyle{fancy}\fancyhead[R]{\thepage}
link-citations: yes
fontsize: 11pt
bibliography: reviewbib.bib
yolo: yes
---

@@

<!-- Structure of the paper -->
The paper is structured as follows. The next section describes the choropleth map, which is the common approach to disease maps and presents examples of atlases in use today. Section 3 discusses the limitations of the choropleth map. Section 4 describes alternative displays, including the cartogram, which is useful when the map has heterogeneously sized geographic units. Section 5 presents the limitations of the production and use of alternative displays. Disease maps are more useful when made interactive, and common options are described in Section 6, along with a discussion of benefits and disadvantages.

@@

# 3 Contemporary alternatives to choropleth maps {#ch:alternatives}

@@

There are many issues to consider when creating alternative map displays: the intended audience of the map, and its purpose are key points in cartogram use and creation as 'there is no "best" cartogram or method of creating cartograms just as there is no "best" map' (Monmonier and Schnell, 1988) [@ACTUC]. Nusrat and Kobourov [@SAIC] provided a framework to investigate implementations of the many algorithms presented, and the "statistical accuracy, geographical accuracy, and topological accuracy".

@@

The white space presents the meaningful empty-space property [@ECGC] [@NAC] but it also distracts the reader from the data, with a low data density [@tufte].

@@

## 3.2 Tile Map
<!-- 173 words -->
A tile map provides a tessellated display of consistent shapes.  A similar method to a rectangular cartogram, represents each geographic area using a square. The squares are tessellated to create a grid. Each area is represented by a square of the same dimensions, each tile is usually one unit of measurement. This could be geographic regions such as states or population-based that use a consistent measure of population for each tile. Regions with more than four neighbors introduce some displacement, because only four squares can be adjoining. Like the chorpleth and cartograms, the tile map uses color to represent a value of a statistic for each area. There are online media sources using this method, which include [NPR](https://www.npr.org/2016/10/18/498406765/npr-battleground-map-hillary-clinton-is-winning-and-its-not-close), [FiveThirtyEight](https://projects.fivethirtyeight.com/2016-election-forecast/), [The Wall Street Journal](http://graphics.wsj.com/elections/2016/2016-electoral-college-map-predictions/), [The Washington Post](https://www.washingtonpost.com/graphics/politics/2016-election/50-state-poll/). Tile maps often need to be created manually, requiring more time and care as the number of geographic areas increases.
<!-- cite mosaic cartograms -->

@@

# 4 Comparison and critique of alternative displays {#sec:compare}
<!-- 408 words -->

@@

# 5 User interaction {#ch:interacting}
<!-- 634 words -->

>>>>

---
title: "Cancer Applications of Choropleth Maps, and the Potential of Cartograms and Alternative Map Displays"
author: "Stephanie Kobakian^[Science and Engineering Faculty, Queensland University of Technology.], Dianne Cook^[Faculty of Econometrics and Business Statistics, Monash University.] and Jessie Roberts^[Queensland University of Technology.]"
output:
  word_document:
    toc: no
    toc_depth: '3'
  pdf_document:
    dev: jpeg
    fig_caption: yes
    number_sections: yes
    toc: no
    toc_depth: 3
geometry: margin = 1.2in
header-includes:
    - \usepackage{setspace}\doublespacing
link-citations: yes
csl: springer-vancouver.csl
fontsize: 11pt
bibliography: reviewbib.bib
yolo: yes
---

@@

<!-- Structure of the paper -->
The paper is structured as follows. The next section describes the choropleth map, which is the common approach to disease maps and presents examples of atlases in use today and discusses the limitations of the choropleth map. Section 3 describes alternative displays, including the cartogram, which is useful when the map has heterogeneously sized geographic units. Section 4 presents the limitations of the production and use of alternative displays. Disease maps are more useful when made interactive, and common options are described in Section 5, along with a discussion of benefits and disadvantages.

@@

# 3. Contemporary alternatives to choropleth maps {#ch:alternatives}

@@

There are two key issues to consider when creating alternative map displays, (1) the intended audience of the map, and (2) its purpose. Nusrat and Kobourov [@SAIC] provided a framework to investigate implementations of the many algorithms presented, and the "statistical accuracy, geographical accuracy, and topological accuracy".

@@

The white space presents the meaningful empty-space property [@ECGC] [@NAC] but it also distracts the reader from the data, with a low data density [@TVDQI].

@@

## 3.2 Tile Map
<!-- 173 words -->
A tile map provides a tessellated display of consistent shapes.  A similar method to a rectangular cartogram, represents each geographic area using a square. The squares are tessellated to create a grid. Each area is represented by a square of the same dimensions, each tile is usually one unit of measurement, this could be geographic regions such as states or population-based that use a consistent measure of population for each tile. Regions with over four neighbors require some necessary displacement. The tile map uses color to represent a value of a statistic for each area. A similar method to a rectangular cartogram represents each geographic area using a square of the same dimensions. There are online media sources using this method, these include [@NPR], [@FiveThirtyEight], [@WSJ], [@WP]. Tile maps may be difficult to create as they are best created manually, they require additional time and care as the number of geographic areas to include increases.

@@

# 4. Comparison and critique of alternative displays {#sec:compare}
<!-- 408 words -->

@@

# 5. User interaction {#ch:interacting}
<!-- 634 words -->
